---
title: "Bioinformatics"
author: "Erik Larsen"
date: "7/17/2021"
output:
  html_document:
    toc: yes
    toc_float:
      collapsed: yes
    numbered_sections: yes
    code_folding: hide
---

# Overview

+ I developed this Markdown to document some bioinformatics analyses of data published by **M. Bhattacharya Lab** in the publication, [Transmembrane protein TMEM184B is necessary for interleukin-31–induced itch](https://journals.lww.com/pain/Abstract/9000/Transmembrane_protein_TMEM184B_is_necessary_for.97918.aspx).

+ These analyses were performed on a number of different datasets and include:

  + MA plots
  + volcano plots
  + heatmaps
  + bar graphs 
  
1. `Illumina FASTQ files` were pseudoaligned  with the `Salmon` algorithm

2. run through `Usegalaxy.org`'s `DESeq2` wrapper algorithm to obtain `DGEA files`

3. `Galaxy` also returned TPM data for use in transcriptional profiling (heatmaps)

+ To obtain pathway and gene ontology info, the [Enrichr](https://maayanlab.cloud/Enrichr/) database was queried from the console using the [enrichR](https://github.com/guokai8/Enrichr#:~:text=Description%20EnrichR%20is%20a%20package%20can%20be%20used,species%20pubished%20by%20ENSEMBL%20and%20included%20with%20Bioconductor.) package.

+ The source script of this file is the [Bioinformatics R Script](C:/Users/Erik/Desktop/BoxCopy/Programming Scripts and Data/Bio/Scripts/R/Broad/Bioinformatics Script.R).  

+ Standalone analyses or tutorials, including volcano plotting, and `gene ontology` and `pathway analysis` can be found in other `R/Github Markdown` files and scripts, [Gene Ontology Analysis Snippet](https://github.com/eriklarsen4/Proteomics/blob/main/GO%20Analysis/GO-Analysis.md) and
[Volcano plot tutorial R Markdown](https://github.com/eriklarsen4/ggplot-scripts/blob/master/Bioinformatics/Volcano-plot-tutorial.md)

# Environment Prep {.tabset .tabset-pills .tabset-fade}

Note that code can be sectioned and condensed with the `Alt + O` command.

Packages for this script:
[tidyverse](https://cran.r-project.org/package=tidyverse), [stringr](https://cran.r-project.org/package=stringr),
[readr](https://cran.r-project.org/package=readr), [biomaRt](https://bioconductor.org/packages/biomaRt/),
[GO.db](http://bioconductor.riken.jp/packages/3.0/data/annotation/html/GO.db.html), [PANTHER.db](https://bioconductor.org/packages/PANTHER.db/), [BiocGenerics](https://bioconductor.org/packages/BiocGenerics),
[pheatmap](https://cran.r-project.org/package=pheatmap)

## Install and Load Packages

Install biology-based packages with `BiocManager`.

Load the [BiocGenerics](https://bioconductor.org/packages/BiocGenerics) package for Bioconductor-relevant functionality (installing packages from [Bioconductor](https://www.bioconductor.org/))

```{r Install 1, include = T, message = F, warning = F, echo = T}
library(BiocGenerics) ## Needed to install and/or load Bioconductor packages
```

``` {r Install 2, include = F, message = F, warning = F, echo = F}
library(PANTHER.db) ## Needed in the Enrichr package when querying pathways
pthOrganisms(PANTHER.db) = "MOUSE"
```

Load the remaining packages

``` {r Install 3, include = T, message = F, warning = F, echo = T}
library(enrichR) ## Taps the Enrichr database; much better utility than the website

  ## For data wrangling (slicing/adding/removing/melting/rearranging dataframes and their columns and rows):
library(tidyverse)
# library(plyr)
# library(dplyr)
# library(reshape2)

library(readr) ## For importing data and files

library(stringr) ## Awesome for manipulating strings

  ## Databases for downstream analyses of lists via gene ontology:
library(biomaRt)
#library(org.Hs.eg.db)
library(org.Mm.eg.db)
library(GO.db)

library(pheatmap) ## For creating awesome heatmaps

library(httpuv) ## For including plots in Markdown file output
```

## Upload Enrichr DBs

Obtain databases from `enrichr` by creating a variable and viewing it. Subset from the variable a few databases of interest.

```{r Upload Enrichr db, include = T, message = F, warning = F, echo = T}
    ## Store it as a variable for better visualization and eventual subsetting
DBs = listEnrichrDbs()

  ## View the variable in the editor to find the relevant databases and their indeces for subsetting
  ## (click on the dataframe in the Global Environment and manually peruse)

diff_DBs = DBs %>%
  dplyr::filter(grepl(libraryName, pattern = "Transcription_Factor_PPIs|Reactome_2022|Most_Popular_Genes|ENCODE_TF|GPCR|Enrichr")) %>%
  dplyr::select(libraryName) %>%
  as.vector() %>%
  unlist() %>%
  as.character()

DBs = DBs %>%
  dplyr::filter(grepl(libraryName, pattern = "GO_.+?(2023)$|Panther_2016|PPI_Hub|Jensen_DISEASES")) %>%
  dplyr::select(libraryName) %>%
  as.vector() %>%
  unlist() %>%
  as.character()
```


```{r old Enrichr, include = F, message = F, warning = F, echo = F}
# DBs = listEnrichrDbs() ## Store it as a variable for better visualization and eventual subsetting
# diff_DBs = DBs$libraryName[c(50,127,91,92,105,162)]
# DBs = DBs$libraryName[c(130,131,132,102,14,110)]
# 
#   ## Peek at the selections
# diff_DBs
# DBs

```

Create dataframes to house bioinformatics analyses returned from enrichr.

``` {r include = TRUE, message = FALSE}
  ## Create dataframes to store returned analyses as variables
GO_Processes = data.frame()
GO_Cell_Comps = data.frame()
GO_Mol_Funcs = data.frame()
PATHWAYS = data.frame()
PPIs = data.frame()
DISEASES = data.frame()
ENCODE_TFs = data.frame()
Enrichr_TFs = data.frame()
GPCR_Down_Perturbs = data.frame()
GPCR_Up_Perturbs = data.frame()
GO_PROCESS_GENES = data.frame()
GO_MOL_FUNC_GENES = data.frame()
GO_CELL_COMP_GENES = data.frame()
PATHWAY_GENES = data.frame()
PPI_GENES = data.frame()
DISEASE_GENES = data.frame()
ENCODE_TF_GENES = data.frame()
ENRICHR_TF_GENES = data.frame()
GPCR_DOWN_PERTURB_GENES = data.frame()
GPCR_UP_PERTURB_GENES = data.frame()
```

## Source Enrichr Function

Create a data frame to collect data returned by `enrichr`; then create the function that will tap enrichr's database and populate all the dataframes.

``` {r Old Enrichr fn, include = F, message = F, warning = F, echo = F}
  ## Create a function that will return Enrichr-based functional analysis on a given gene set
    ## First, create a data frame that will house combined enrichr output
      ## (this is for exporting collapsed, collective data into a CSV)
Enrichr_Functional_Output = data.frame()

  ## Create the analysis function
Enrichr_Analysis = function(Genes){
  ## first: Gene ontology biological processes; this identifies cell/biological processes with which the genes in the dataset are associated
  GO_Processes = as.data.frame(
    enrichr(
      c(Genes), DBs[1])
  )
  
  
  ## Remove unnecessary columns
  GO_Processes = GO_Processes[,-c(5,6,7)]
   
  
  ## Clean up the dataframes
  GO_Processes = GO_Processes %>%
    separate(GO_Biological_Process_2018.Overlap, c("Overlap", "Process Size"), "/")
  ## Convert the values into integers for computation
  GO_Processes$Overlap = as.numeric(GO_Processes$Overlap)
  GO_Processes$`Process Size` = as.numeric(GO_Processes$`Process Size`)
  
  
  ## Add Enrichr's Z-score. This determines a term/pathway's rank change across
  ## all terms/pathways when compared to the term/pathway's "pre-rank" based on
  ## multiple simulations of the same number of enriched genes.
  GO_Processes = add_column(GO_Processes, EnrichrZscore = GO_Processes$GO_Biological_Process_2018.Combined.Score/log(GO_Processes$GO_Biological_Process_2018.P.value), .before = 6)
  ## Add the "weighted overlap ratio" of a term/pathway's number of enriched genes. 
    ## The fraction of differentially expressed genes belonging to a term/pathway
    ## ("overlap"; numerator) out of the total genes in that term/pathway/hub (denominator)
    ## is scaled by the overlap. This weights differentially expressed genes by pathway size.
  GO_Processes = add_column(GO_Processes, Weighted_Overlap_Ratio = GO_Processes$Overlap*(GO_Processes$Overlap/GO_Processes$`Process Size`), .before = 4)
  ## Add the modifcation of Enrichr's "Combined Score". This scales the term/pathway's Adjusted P-value by the Enrichr Z-score. 
  GO_Processes = add_column(GO_Processes, Modified_Combined_Score = (abs(GO_Processes$EnrichrZscore))*(-log(GO_Processes$GO_Biological_Process_2018.Adjusted.P.value)), .before = 6)
  ## Remove GO IDs
  GO_Processes$GO_Biological_Process_2018.Term = GO_Processes$GO_Biological_Process_2018.Term %>% gsub(x = GO_Processes$GO_Biological_Process_2018.Term, pattern = " \\(.*\\)$", replacement = "")
  
 
  ## Extract a concatenated string containing genes involved in each GO Biological Process
  GO_PROCESS_GENES = str_split(GO_Processes$GO_Biological_Process_2018.Genes, pattern = ";", simplify = TRUE)
  for (i in 1:nrow(GO_PROCESS_GENES)){
    GO_PROCESS_GENES[i,] = paste(substr(GO_PROCESS_GENES[i,], 1, 1),
                                 tolower(substr(GO_PROCESS_GENES[i,], 2, 7)), sep = "") 
  }
  
  
  ## Make sure the Enrichr reference results are correctly ordered
  GO_Processes = GO_Processes[order(GO_Processes$Modified_Combined_Score, decreasing = TRUE), ]
  ## Re-set the index to make sure indeces are correct
  row.names(GO_Processes) = NULL
  
  
  ## Export these data frames to the global environment
  GO_Processes <<- GO_Processes
  
  return(c("GO BIOLOGICAL PROCESSES", GO_Processes[1:10,1]))
}
```

```{r Enrichr Function, include = T, message = F, warning = F, echo = T}
source("~/GitHub/Proteomics/Functions/Enrichr Analysis Function.R")
```

## Data import

Import the adult `DRG DESeq2 file`

``` {r DESeq2 import, include = T, message = F, warning = F, echo = T}
aDRG = read.csv("~/GitHub/ggplot-scripts/Bioinformatics/RNAseq Data Files/DESeq2 Expression Results.csv")

  ## Filter (subset) genes that went undetected or were outliers in terms of counts;
  ## new dataframe should not contain any NAs in p-value columns
aDRG3 = subset(aDRG, (!is.na(aDRG[,"AdjP"])))
 ## Filter the DEGs by removing rRNAs and mitochondrial tRNAs, along with pseudogenes, etc.
aDRG9 = aDRG3 %>% filter(!grepl(GeneID,
                                pattern = "Rps.+.?$|RP.+.?$|Rpl.+.?$|MRPL.+.?$|Mrpl.+.?$|MRPS.+.?$|Mrps.+.?$|.*Rik.+$|.*Rik$|Gm.+.?$|^[A-Z]+[A-Z].+.?$|^[0-9]+.+.?$"))
#mt.+.?$|  <-- string identifier for mitochondrial tRNAs

  ## Create a column in the DESeq2 dataframe that scales the Adjusted P-value by log-base 10
aDRG9$log10ADJP = -log10(aDRG9$AdjP)

  ## Add a column to the DEG dataset that contains a string, describing whether the gene is differentially expressed
    ## First create the column and use Gene IDs as place-holders
aDRG9$g.o.i. = aDRG9$GeneID
  ## Replace DEGs with the string, "DEGs"
aDRG9$g.o.i.[which(aDRG9$AdjP <= 0.05)]= "DEGs"
  ## Replace the remaining genes with "Non-DEGs"
aDRG9$g.o.i.[which(aDRG9$AdjP >= 0.05)]= "Non-DEGs"

aDRG_DEG_list = c(aDRG9$GeneID[which(aDRG9$g.o.i. == "DEGs")])

```

``` {r e13, include = F, message = F, warning = F, echo = F}
e13_DRG = read.csv("C:/Users/Erik/Desktop/BoxCopy/Lab/Omics/RNAseq/Embryonic DRG/Processed Galaxy Output/Test Results to Upload/e13/DESeq2_result_file_on_GT_e13_PARA.csv", header = FALSE)

  ## Set the column names
colnames(e13_DRG) = colnames(aDRG)[c(1:7)]
  ## Subset the data as with the adult dataset
e13_DRG3 = subset(e13_DRG, (!is.na(e13_DRG[,"AdjP"])))
e13_DRG9 = e13_DRG3 %>% filter(!grepl(e13_DRG3$GeneID, pattern = "Rps.+.?$|RP.+.?$|Rpl.+.?$|MRPL.+.?$|Mrpl.+.?$|MRPS.+.?$|Mrps.+.?$|.*Rik.+$|.*Rik$|Gm.+.?$|^[A-Z]+[A-Z].+.?$|^[0-9]+.+.?$"))

  ## Add a column containing "%WT"
e13_DRG9 = add_column(e13_DRG9, X..WT = 2^(e13_DRG9$log2.FC.[]))
  ## Rename the column
colnames(e13_DRG9)[8] = colnames(aDRG)[8]
  ## Convert the values for graphics (positive from 0 -> Inf)
e13_DRG9$log10ADJP = -log10(e13_DRG9$AdjP)


e13_DRG9$g.o.i. = e13_DRG9$GeneID
e13_DRG9$g.o.i.[which(e13_DRG9$AdjP <= 0.01)]= "DEGs"
e13_DRG9$g.o.i.[which(e13_DRG9$AdjP >= 0.01)]= "Non-DEGs"

e13_DRG_DEG_list = c(e13_DRG9$GeneID[which(e13_DRG9$g.o.i. == "DEGs")])

```

```{r p0, eval = F, include = F, message = F, warning = F, echo = F}
P0_DRG = read.csv("C:/Users/Erik/Desktop/BoxCopy/Lab/Omics/RNAseq/Embryonic DRG/Processed Galaxy Output/Test Results to Upload/p0/DESeq2_result_file_on_GT_p0_PARA.csv", header = FALSE)

colnames(P0_DRG) = colnames(aDRG)[c(1:7)]

P0_DRG3 = subset(P0_DRG, (!is.na(P0_DRG[,"AdjP"])))
P0_DRG9 = P0_DRG3 %>% filter(!grepl(P0_DRG3$GeneID, pattern = "Rps.+.?$|RP.+.?$|Rpl.+.?$|MRPL.+.?$|Mrpl.+.?$|MRPS.+.?$|Mrps.+.?$|.*Rik.+$|.*Rik$|Gm.+.?$|^[A-Z]+[A-Z].+.?$|^[0-9]+.+.?$"))


P0_DRG9 = add_column(P0_DRG9, X..WT = 2^(P0_DRG9$log2.FC.[]))
colnames(P0_DRG9)[8] = colnames(aDRG)[8]

P0_DRG9$log10ADJP = -log10(P0_DRG9$AdjP)

P0_DRG9$g.o.i. = P0_DRG9$GeneID
P0_DRG9$g.o.i.[which(P0_DRG9$AdjP <= 0.05)]= "DEGs"
P0_DRG9$g.o.i.[which(P0_DRG9$AdjP >= 0.05)]= "Non-DEGs"

P0_DRG_DEG_list = c(P0_DRG9$GeneID[which(P0_DRG9$g.o.i. == "DEGs")])

```

``` {r p10, eval = F, include = F, message = F, warning = F, echo = F}
P10_DRG = read.csv("C:/Users/Erik/Desktop/BoxCopy/Lab/Omics/RNAseq/Embryonic DRG/Processed Galaxy Output/Test Results to Upload/p10/DESeq2_result_file_on_GT_p10_PARA.csv", header = FALSE)

colnames(P10_DRG) = colnames(aDRG)[c(1:7)]


P10_DRG3 = subset(P10_DRG, (!is.na(P10_DRG[,"AdjP"])))
P10_DRG9 = P10_DRG3 %>% filter(!grepl(P10_DRG3$GeneID, pattern = "Rps.+.?$|RP.+.?$|Rpl.+.?$|MRPL.+.?$|Mrpl.+.?$|MRPS.+.?$|Mrps.+.?$|.*Rik.+$|.*Rik$|Gm.+.?$|^[A-Z]+[A-Z].+.?$|^[0-9]+.+.?$"))

P10_DRG9 = add_column(P10_DRG9, X..WT = 2^(P10_DRG9$log2.FC.[]))
colnames(P10_DRG9)[8] = colnames(aDRG)[8]


P10_DRG9$log10ADJP = -log10(P10_DRG9$AdjP)

P10_DRG9$g.o.i. = P10_DRG9$GeneID
P10_DRG9$g.o.i.[which(P10_DRG9$AdjP <= 0.05)]= "DEGs"
P10_DRG9$g.o.i.[which(P10_DRG9$AdjP >= 0.05)]= "Non-DEGs"

P10_DRG_DEG_list = c(P10_DRG9$GeneID[which(P10_DRG9$g.o.i. == "DEGs")])

```

``` {r IV data, eval = F, include = F, message = F, warning = F, echo = F}
DIV14 = read.csv("C:/Users/Erik/Desktop/BoxCopy/Lab/Omics/RNAseq/Embryonic DRG/Processed Galaxy Output/Test Results to Upload/Cultured Embryos for Genotype at DIV14.csv")

colnames(DIV14) = colnames(aDRG)[c(1:7)]

DIV143 = subset(DIV14, (!is.na(DIV14[,"AdjP"])))
DIV149 = DIV143 %>% filter(!grepl(DIV143$GeneID, pattern = "Rps.+.?$|RP.+.?$|Rpl.+.?$|MRPL.+.?$|Mrpl.+.?$|MRPS.+.?$|Mrps.+.?$|.*Rik.+$|.*Rik$|Gm.+.?$|^[A-Z]+[A-Z].+.?$|^[0-9]+.+.?$"))

DIV149 = add_column(DIV149, X..WT = 2^(DIV149$log2.FC.[]))
colnames(DIV149)[8] = colnames(aDRG)[8]

DIV149$log10ADJP = -log10(DIV149$AdjP)

DIV149$g.o.i. = DIV149$GeneID
DIV149$g.o.i.[which(c(DIV149$AdjP < 0.01 & DIV149$log2.FC. > 0))]= "Up DEGs"
DIV149$g.o.i.[which(c(DIV149$AdjP < 0.01 & DIV149$log2.FC. < 0))]= "Down DEGs"
```


# Query the Enrichr db {.tabset .tabset-pills .tabset-fade}

Search the `Enrichr` database using the function created above with any of the above gene lists.

```{r Run enrichr snippet, include = T, message = F, warning = F, echo = T}
Enrichr_Analysis(Genes = aDRG_DEG_list)
```

## Prep data for heatmaps

Import the adult transcriptional profile (normalized TPM).

``` {r Trx profile upload, include = T, warnings = F, message = F, echo = T}
  ## Import the adult normalized counts file.
    ## These are expression estimates for each gene, for each sample/replicate,
    ## where each gene's value is normalized to its sample's effect size
aTPM = read_csv("~/GitHub/ggplot-scripts/Bioinformatics/RNAseq Data Files/RNASeqRepResults.csv", col_names = TRUE)
  ## Rename columns
colnames(aTPM) = c("GeneID", "WT1", "WT2", "WT3", "WT4", "Mut1", "Mut2", "Mut3", "Mut4")

  ## Subset by only genes in the filtered aDRG DESeq2 file
ExpProfile = aTPM[aTPM$GeneID %in% c(aDRG9$GeneID),]
```

Create a function that makes a dataframe housing the Z-scored transcriptional profile to then graph using `pheatmap`.

```{r Zscore Function, include = T, warning = F, message = F, echo = F}
  ## Create a function that determines the Zscores of the expression profile (row Z-scores)
Find_Row_Z = function(Expression_Profile){
  ## Create dataframes to calculate Z-scores across replicates of each gene.
  ## These will hold the original data frame values until filled in later commands
  ## "Big" will store the means and sds of each gene
  NetBIG = Expression_Profile
  ## "3" will store only the means by genes for all replicates
  Net3 = Expression_Profile
  ## "4" will store only the sds by gene for all replicates
  Net4 = Expression_Profile
  ## "Z" will store only the Z-scores, which will be used directly to create the heatmaps
  NetZ = Expression_Profile
  
  ## Create a new column to fill with the means and standard deviations of each gene's transcript counts per million ("TPM")
  NetBIG$mean = 0
  NetBIG$sd = 0
  
  ## Loop through the dataframe and fill the "mean" and "sd" columns with their appropriate values
  for (i in 1:nrow(NetBIG)){
    NetBIG$mean[i] = (sum(Expression_Profile[i,c(2:ncol(Expression_Profile))])/ncol(Expression_Profile[,c(2:ncol(Expression_Profile))]))
  }
  for (j in 1:nrow(NetBIG)){
    NetBIG$sd[j] = sd(Expression_Profile[j,c(2:ncol(Expression_Profile))], na.rm = TRUE)
  }
  ## Create a dataframe storing the gene-specific mean normalized TPM in all columns/replicates for Z-score calculating
  Net3[,c(2:ncol(Expression_Profile))] = NetBIG$mean
  ## Create a dataframe storing the gene-specific normalized TPM standard deviationsin all columns/replicates for Z-score calculating
  Net4[,c(2:ncol(Expression_Profile))] = NetBIG$sd
  ## Create the Z-score dataframe
  NetZ[,c(2:ncol(Expression_Profile))] = (Expression_Profile[,c(2:ncol(Expression_Profile))] - Net3[,c(2:ncol(Expression_Profile))])/Net4[,c(2:ncol(Expression_Profile))]
  ## Remove the genes that were detected to have 0 TPMs across all samples
  ## Create a function that evaluates a vector/dataframe's (x's) numerical values. Returns equal length vector with T/F bools.
  ## Subset the dataframe that filters those genes.
  row_has_na = apply(NetZ, 1, function(x){any(is.na(x))})
  NetZ = NetZ[!row_has_na,]
  Z <<- NetZ
}
```

Run the Z-score function.

```{r Z score call, include = T, message = F, warning = F, echo = T}
Find_Row_Z(Expression_Profile = ExpProfile)
```

Arrange the data for generating a full transcriptional profile.

```{r Z rearrange, include = T, message = F, warning = F, echo = T}
  ## Don't remove non-DEGs
Z = Z %>%
  filter(GeneID %in% aDRG9$GeneID[])

  ## Create a list of gene names ordered by euclidean distance by which to re-order the dataframe
Euclid_dist_order = hclust(dist(Z[,c(2:9)], method = "euclidean"))$order
  ## The names (not the numbers)
Euclid_dist_ord_Genes = c(Z$GeneID[Euclid_dist_order])

  ## Transform again to order by clusters (Euclidean distances)
Za = Z %>%
  mutate(GeneID =  factor(GeneID, levels = Euclid_dist_ord_Genes)) %>%
  arrange(GeneID)

  ## Find where the "Itch-related DEGs" are in the clustered matrix subsetted to DEGs for the heatmap
itch_DEGs = c("Il31ra", "Cysltr2", "Npy2r", "Sst",
              "Htr1a", "P2rx3", "Lpar3", "Lpar5",
              "Scn11a", "Scn10a", "Mrgprd", "Trpc6",
              "Trpc3", "F2rl2", "Htr1f", "Osmr",
              "Fxyd2", "Htr4", "Mrgprx1", "Ptgdr",
              "Trpa1", "Trpm6", "Hrh1", "Mrgpra3", "Tmem184b")

temp = vector()
for( i in 1:length(itch_DEGs)){
  temp[i] = which(Euclid_dist_ord_Genes == itch_DEGs[i])
}
itch_index = temp
#itch_index

  ## Confirm by indexing; copy and paste to customize the clustered heatmap
#Euclid_dist_ord_Genes[c(itch_index)]
  ## Confirm the location/order of the transformed matrix/dataframe
  ## Should be "Il31ra"
Za[itch_index[1],]

```

# Plot the heatmaps {.tabset .tabset-pills .tabset-fade}

## Adult Full

View the full transcriptional profile of **Adult Tmem184b-mutant DRG neurons**.

```{r Full Z-scored, Euclidean-clustered and ordered pheatmap, include = T, message = F, warning = F, echo = T, render = T}

 ## Visualize the Z-scored, Euclidean-clustered and ordered gene TPMs across replicates with "pheatmap"
pheatmap(mat = Za[,2:9],
         color = colorRampPalette(c("navy", "white", "darkgoldenrod4"))((50)),
         clustsering_distance_rows = "euclidean",
         angle_col = 0,
         cutree_rows = 2,
         treeheight_row = 35,
         treeheight_col = 7,
         show_rownames = F)
```

## Adult Zoom

**View the cluster near Tmem184b**

```{r Tmem184b cluster Zoom on Z-scored Euclidean-clustered and ordered pheatmap, include = T, message = F, warning = F, echo = T, render = T}
  ## Visualize around the Tmem184b cluster
pheatmap(mat = Za[1597:1637,2:9],
         color = colorRampPalette(c("navy", "white", "darkgoldenrod4"))((50)),
         clustsering_distance_rows = "euclidean",
         angle_col = 0,
         treeheight_row = 35,
         treeheight_col = 9,
         labels_row = Euclid_dist_ord_Genes[1597:1637])
```

## Adult Itch 

**View the cluster including select itch transcripts**.

```{r Itch trx cluster Zoom on Z-scored Euclidean-clustered and ordered pheatmap, include = T, message = F, warning = F, echo = T, render = T}
  ## Visualize the "Itch-related DEGs"
pheatmap(mat = Za[c(itch_index),2:9],
         color = colorRampPalette(c("navy", "white", "darkgoldenrod4"))((50)),
         clustsering_distance_rows = "euclidean",
         angle_col = 0,
         treeheight_row = 35,
         treeheight_col = 9,
         cutree_rows = 4,
         labels_row = Euclid_dist_ord_Genes[c(itch_index)])
```

## Adult DEGs Only

Arrange the data to include only DEGs.

``` {r sort, include = T, message = F, warning = F, echo = T}
  ## Transform the profile to include only DEGs
Zb = Z %>%
  filter(GeneID %in% aDRG9$GeneID[c(1:376)])
```

Prep it for a heatmap as before (not shown).

```{r Itch DEG heatmap prep, include = T, warning = F, message = F, echo = F}
  ## Create a list of gene names ordered by euclidean distance by which to re-order the dataframe
Euclid_dist_order = hclust(dist(Zb[,c(2:9)], method = "euclidean"))$order
  ## The names (not the numbers)
Euclid_dist_ord_Genes = c(Zb$GeneID[Euclid_dist_order])

  ## Transform again to order by clusters (Euclidean distances)
Zb = Zb %>%
  mutate(GeneID =  factor(GeneID, levels = Euclid_dist_ord_Genes)) %>%
  arrange(GeneID)

temp = vector()
for( i in 1:length(itch_DEGs)){
  temp[i] = which(Euclid_dist_ord_Genes == itch_DEGs[i])
}
itch_DEG_index = temp
#itch_DEG_index

  ## Confirm by indexing; copy and paste to customize the clustered heatmap
#Euclid_dist_ord_Genes[c(itch_DEG_index)]
  ## Confirm the location/order of the transformed matrix/dataframe
  ## Should be "Il31ra"
#Zb[itch_DEG_index[1],]
```

**Plot the results of the profile of only DEGs in Tmem184b-mutant DRG neurons.**

```{r Z-scored Euclidean-clustered and ordered pheatmap of DEGS, include = T, message = F, warning = F, echo = F, render = T}
  ## Visualize the Z-scored, Euclidean-clustered and ordered DEGs across replicates with "pheatmap"
pheatmap(mat = Zb[,2:9],
         color = colorRampPalette(c("navy", "white", "darkgoldenrod4"))((50)),
         clustsering_distance_rows = "euclidean",
         clustering_distance_cols = "euclidean",
         angle_col = 0,
         treeheight_row = 35,
         treeheight_col = 7,
         show_rownames = F)
```

## Adult DEGs Only Zoom

**Zoom to Tmem184b**

```{r Tmem184b Cluster Zoom Z-scored Euclidean-clustered ordered pheatmap of DEGs, include = T, message = F, warning = F, echo = F, render = T}
  ## Visualize around the Tmem184b cluster
pheatmap(mat = Zb[92:132,2:9],
         color = colorRampPalette(c("navy", "white", "darkgoldenrod4"))((50)),
         clustsering_distance_rows = "euclidean",
         angle_col = 0,
         treeheight_row = 35,
         treeheight_col = 9,
         labels_row = Euclid_dist_ord_Genes[92:132])
```

## Adult DEGs Only Itch Zoom

**Zoom to the area comprising many itch genes**

```{r Itch DEG Cluster Zoom Z-scored Euclidean-clustered ordered pheatmap of DEGs, include = T, message = F, warning = F, echo = F, render = T}
  ## Visualize the "Itch-related DEGs"
pheatmap(mat = Zb[c(87:376),2:9],
         color = colorRampPalette(c("navy", "white", "darkgoldenrod4"))((50)),
         clustsering_distance_rows = "euclidean",
         angle_col = 0,
         treeheight_row = 35,
         treeheight_col = 9,
         cutree_rows = 5,
         show_rownames = F)
```

**View select itch genes**.

+ The difference between the first three heatmaps and the last four heatmaps is hierarchically clustering only on DEGs in the last four

+ Clustering was performed on all genes in the first three

```{r Itch DEG only of Z-scored Euclidean-clustered ordered pheatmap of DEGs, include = T, message = F, warning = F, echo = F, render = T}
  ## Visualize the "Itch-related DEGs"
pheatmap(mat = Zb[c(itch_DEG_index),2:9],
         clustering_distance_rows = "euclidean",
         color = colorRampPalette(c("navy", "white", "darkgoldenrod4"))((50)),
         angle_col = 0,
         treeheight_row = 35,
         treeheight_col = 9,
         cutree_rows = 4,
         labels_row = Euclid_dist_ord_Genes[c(itch_DEG_index)])
```
