---
title: "Gene Ontology Analysis Snippet"
author: "Erik Larsen"
date: "9/17/2021"
output: 
  html_document:
    toc: yes
    toc_float:
      collapsed: yes
    numbered_sections: yes
    code_folding: hide
---

# Overview

This markdown serves as documentation of gene ontology analysis of
multiple datasets generated in the lab of **Dr. Martha Bhattacharya**,
two of which have been used for publications
([here](https://journals.lww.com/pain/abstract/2022/05000/transmembrane_protein_tmem184b_is_necessary_for.18.aspx),
[here](https://www.biorxiv.org/content/10.1101/2022.05.27.493793v3#:~:text=Taken%20together%2C%20our%20data%20suggest%20that%20TMEM184B%20is,be%20linked%20to%20neurodevelopmental%20disorders%20than%20to%20dementia.)).
The totality of the example dataset is private (preliminary, unpublished data).

+ `Gene ontology` and `pathway enrichment analyses` provide high-level analysis of the cell/mol consequences following
perturbations (genetic or pharmacological) of interest. This is incredibly powerful for studying novel proteins.

  + As such, the approach has a few **problems**:
  
    1.  When analyzing a list of genes or proteins quantitatively changed by a perturbation,
        implicated biological processes or pathways depend on a database's curation of data
        and the user's knowledge of biological processes or pathways.

    2.  Genes and proteins have aliases; these can cause discrepancies
        missed by researchers querying gene ontologies when their
        supplied list does not contain the “correct” alias
        
        + i.e. `GO Terms` may be associated with some genes/proteins
          but not all aliases

    3.  Some genes/proteins may have been mistakenly omitted from a
        `GO Term`

    4.  It's probably useful to understand putative connections between a
        given gene/protein & its gene ontologies with *other*
        genes/proteins & *their* gene ontologies— in particular,
        accounting for as many aliases as possible

        -   If a user is interested in a manipulation’s effect on cellular
            processes, this approach may help better connect otherwise
            unconnected genes/proteins to other pathways/processes
  
  + Thus, I created multiple functions, harnessing the `GO database`’s
    repository, but with more flexibility than just plugging and chugging into
    `geneontology.org`'s or `enrichr`'s websites
    
  + This snippet demonstrates the creation and use of one particular
    function (another that addresses issues is documented [here]()):

    + `GO_INFO_fn`, for performing the statistical analyses of the
      genes/proteins in a provided list
      
      + More specific questions and their follow-up functions to the `GO_INFO_fn`
        are specified in the `Additional Functions` tab of the `Function definitions` section below

# Function definitions {.tabset .tabset-pills .tabset-fade}

## GO_INFO_fn

+ For the GO and pathway over-representation analysis of a list of interest (e.g. analysis of differentially expressed genes)

+ The `GO_INFO_fn` takes, as input:
    +   a list of gene or protein IDs
    +   the species of the list

+ it returns to the Global Environment:
  
  + a dataframe, `GO_INFO_by_GENE_df`, that houses:
    + each gene/protein from a provided list (the input list)
      + all the `# of GO Terms` to which a gene/protein has been annotated by the `GO.db`
      + all the `GO IDs` to which a gene/protein has been annotated by the `GO.db`
      + all the `GO Terms` to which a gene/protein has been annotated by the `GO.db`
      
  + a list (`Unique_GOs`) of all the unique `GO Terms` associated with the entire list
  
  + the supplied list of interest, removing genes/proteins that do
    not have terms associated with it ( = `list_of_interest`)

## Additional Functions
The `GO_INFO_by_GENE_df` houses information that can then be used for more targeted questions,
using other functions, such as:

+ Are there any enriched `GO Terms` from the supplied list?

  + if yes, what are they? ( use
    `Find_Enriched_GOs_from_GO_INFO_df_fn`)

  + Which genes/proteins from the list of interest are associated with a
    specific `GO Term`? ( use `Find_Genes_With_X_GO_Term_fn`)

  + Which `GO Terms` are associated with a specific gene/protein from
    the supplied list? ( use `Find_GOs_of_Gene_X_fn`)

  + What are the `GO Terms` of two genes/proteins of from the supplied
    list? ( use `Find_GOs_of_Two_Genes_fn`)

  + Which `GO Terms` are shared by all genes/proteins from the supplied
    list? ( use `Find_Genes_Related_By_GO_Term_fn`)

# Example Environment Prep {.tabset .tabset-pills .tabset-fade}

## Load packages and data

Load packages

```{r Load Packages, include = T, warning = F, message = F, echo = T}
library(stringr)
library(ggplot2)
library(ggrepel)
# library(enrichR)
library(org.Mm.eg.db)
library(org.Hs.eg.db)
library(GO.db)
library(tidyverse)
# library(tidyr)
library(ggpubr)
library(mgcv)
library(splines)
```

Import data (not shown)

```{r Import data, include = T, warning = F, message = F, echo = F}
All_DF_annot = read_csv("C:/Users/Erik/Desktop/BoxCopy/Lab/Omics/MS/All_DF_annot.csv")

whittled_candidates_list = All_DF_annot %>%
  dplyr::filter(p.o.i. == "Hits") %>%
  dplyr::select(HumanGeneID) %>%
  as.vector() %>%
  unlist() %>%
  as.character()

rm(list = ls()[which(grepl(ls(), pattern = "All_DF_annot|whittled_candidates_list") == F)])

All_DF_annot = All_DF_annot %>%
  dplyr::mutate(p.o.i. = case_when(HumanGeneID == "TMEM184B" ~ "Target Bait",
                                   HumanGeneID == "GFP" ~ "Control Bait",
                                   HumanGeneID == "birA" ~ "Control Bait",
                                   grepl(p.o.i., pattern = "Hits") ~ "Hits",
                                   TRUE ~ "Background"),
                labs = case_when(!is.na(labs) & HumanGeneID == "TMEM184B" ~ "Target Bait",
                                 !is.na(labs) & HumanGeneID != "TMEM184B" ~ "Control Bait",
                                 TRUE ~ ""),
                Targetavg = TMEMavg) %>%
  dplyr::select(-1)
```

Source functions

```{r Source Functions, include = T, warning = F, message = F, echo = T}
source("~/GitHub/Proteomics/Functions/GO Analysis Functions.R")
source("~/GitHub/Proteomics/Functions/IPMS Functions.R")
```

## Data Analysis

1. Wrangle the example proteomics data (not shown)

2. Pass the data-wrangled output into an analysis function that narrows the list of proteins of interest (not shown; [snippet here](https://github.com/eriklarsen4/Proteomics/blob/go-analysis-only-rev/IPMS%20Data%20Wrangling%20and%20Analysis/IPMS%20Data%20Processing.md) is still under revision)

3. Filter the dataset. The example list of proteins for this snippet are derived here

```{r Filter candidates, include = T, message = F, warning = F, echo = T}
  ## Remove "RNA polymerase" proteins, translation-related proteins, heat shock proteins, histone proteins, tRNA proteins
All_DF_annot = All_DF_annot %>% 
  as_tibble() %>%
  filter(
    !grepl(
      `Identified Proteins (6388)`,
      pattern = "RNA polymerase|Eukaryotic initiation|Eukaryotic elongation|Heat shock|Histone|tRNA|",
      fixed = TRUE
      )
    )

  ## Sort by descending FC
All_DF_annot = All_DF_annot %>%
  dplyr::arrange(desc(FC))
  ## Re-order the index
row.names(All_DF_annot) = NULL
```

## Visualize Dataset

The "`Hits`" in blue below will be used as the example list of proteins for the `GO_INFO_fn`

```{r Generate ggplot object, include = T, warning = F, message = F, echo = F}
  ## Create the plot. Save it as a variable; add appropriate layers
All_FC = ggplot(data = All_DF_annot) +
  geom_point(data = subset(All_DF_annot, p.o.i. == "Background"),
             aes(x = Controlavg,
                 y = Targetavg,
                 color = p.o.i.),
             alpha = 0.1) +
  geom_point(data = subset(All_DF_annot, p.o.i. == "Hits"),
             aes(x = Controlavg,
                 y = Targetavg,
                 color = p.o.i.),
             alpha = 0.5) +
  geom_point(data = subset(All_DF_annot, p.o.i. == "Control Bait"),
             aes(x = Controlavg,
                 y = Targetavg,
                 color = p.o.i.),
             alpha = 1.0) +
  geom_point(data = subset(All_DF_annot, p.o.i. == "Target Bait"),
             aes(x = Controlavg,
                 y = Targetavg,
                 color = p.o.i.),
             alpha = 1.0) +
  coord_cartesian(xlim = c(0,7.5), ylim = c(0,7.5)) +
  labs(x = "Avg. TSCs in Control IP",
       y = "Avg. TSCs in Target IP",
       title = "IP-MS",
       col = "Protein\nCategory") +
  scale_color_manual(values = c("gray48",
                                "darkgoldenrod3",
                                "navy",
                                "darkgoldenrod4")) +
  theme_bw() +
  
  geom_abline(intercept = 0, slope = 2, linetype = "dashed", color = "firebrick") +
  geom_text(x = 3, y = 7.5, label = "Fold Change \u2265 2", color = "firebrick", size = 4) +
  theme(panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5, size = 20),
        legend.position = c(0.85,0.5),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 15),
        legend.text = element_text(size = 14),
        legend.title = element_blank())

  ## Overwrite to add labels
All_FC = All_FC  + geom_text_repel(data = All_DF_annot,
                                   x = All_DF_annot$Controlavg,
                                   y = All_DF_annot$Targetavg,
                                   color = "black",
                                   aes(label = labs),
                                   size = 4, max.overlaps = Inf)
```

```{r Scatter plot of potential hits, include = T, warning = F, message = F, echo = F, render = T}
All_FC
```

# GO Analysis on List of Interest {.tabset .tabset-pills .tabset-fade}

Conduct GO Analysis on the example list above using the `GO_INFO_fn`:

-   queries the `GO.db` and extracts `Terms` and `IDs` associated with genes/proteins in a provided list

-   output is `GENE_GO_INFO_df` (gene-wise GO information) and
    `GO_INFO_by_TERM_df` (term-wise GO information) in the Global Environment

One approach to GO Analysis is by determining `GO Term` or `Pathway` `Over-representation`, meaning:

- finding the probability that a particular `GO Term`/`Pathway` is more likely than chance to be over- or under-represented in the provided list
  
- To find this probability, we first need to:
  
  + determine the probability of observing the genes/proteins from the supplied list in a given `Term` (for all `GO Term`s)
  
  + these probabilities ("`Overlap_Prob`s") are determined using **Bayes' Theorem**, where:
        
    - `Overlap_Prob` = n<sub>genes</sub> from `list of interest` in the
      given `GO Term` / total_number<sub>genes</sub> in that `GO Term`

  + "`Weighted_Overlap_Prob`s" are computed in an effort to control for list size, heavily annotated genes, and term size, where:

    - `Weighted_Overlap_Prob` = ( (n<sub>genes</sub> from
    `list of interest` in the given `GO Term`) * (`GO_Term_rate`) *
    (`Gene_freq` ) ) / (`list length` / `GO Term size`)

    -   `GO_Term_rate` = n<sub>unique_GO_Terms</sub> from
        `list of interest` / n<sub>total</sub> `GO`s

        - (accounts for the number of `GO Terms` related to the provided list)

    -   `Gene_freq` = sum( P<sub>r</sub> (each gene from
        `list of interest` arising in *any* `GO Term`) )

        - (accounts for heavily annotated genes i.e. **how often does gene X get annotated to a different GO Term?**)

    -   `GO Term size` = n<sub>total_genes</sub> in a `GO Term`

- Finally, a binomial exact test is conducted for each `GO Term` to assess how likely the observed probability of overlap is when compared to a theoretical probability

## GO_INFO_fn

```{r Pass the IPMS list into the GO_INFO_fn, include = T, message = F, warning = F, echo = T}
GO_INFO_fn(list_of_interest = whittled_candidates_list,
           species = "human")
```

## GO Overlap

Find the `Overlap_Prob`s for the list of interest

+ To determine these probabilities ("`Overlap_Prob`s"), use **Bayes' Theorem**
  
+ Assess their distributions, make corrections

```{r v2 GO OverRep Statistical Analyses, include = T, warning = F, message = F, echo = T}

  ## Store GO term information
for (i in 1:nrow(GO_INFO_by_TERM_df) ) {
  
  x = paste(
    unique(
      unlist(
        str_split(as.vector(GO_INFO_by_TERM_df[i, 4]),
                  pattern = ";")))
    , collapse = ";")
  
  GO_INFO_by_TERM_df[i,3] = length(c(unique(unlist(str_split(as.vector(GO_INFO_by_TERM_df[i, 4]),
                                                             pattern = ";")
                                                   )
                                            )
                                     )
                                   )
  GO_INFO_by_TERM_df[i,4] = x
  
}

  ## Store conditional probabilities
GO_INFO_by_TERM_df = GO_INFO_by_TERM_df %>%
  dplyr::mutate(Num_Genes_in_Term = as.numeric(Num_Genes),
                
                Num_Genes_in_Term_from_List = as.numeric(Num_Genes_from_List),
                
                Num_Genes_not_in_Term = max(Num_Genes_in_Term) - Num_Genes_in_Term,
                
                Num_Genes_not_in_Term_from_List = as.numeric(length(list_of_interest)) - Num_Genes_in_Term_from_List,
                
                Num_Genes_in_Term_not_from_List = Num_Genes_in_Term - Num_Genes_in_Term_from_List,
                
                Num_Genes_not_from_List = max(Num_Genes_in_Term) - as.numeric(length(list_of_interest)),
                
                Num_Genes_not_in_Term_not_from_List = Num_Genes_not_from_List - Num_Genes_in_Term_not_from_List) %>%
  
  dplyr::select(-Num_Genes, -Num_Genes_from_List) %>%
  dplyr::arrange(desc(Num_Genes_in_Term))
```

To find over-representation, determine the probabilities that genes from the `list of interest` are found in a given `GO Term`:

+ `Overlap_Prob` = n~genes~ from `list of interest` in the given `GO Term` / tot_n~genes~ in that `GO Term`

+ `Weighted_Overlap_Prob` = ( (n~genes~ from `list of interest` in the given `GO Term`) * (`GO_Term_rate`) * (`Gene_freq` ) ) / (`list length` / `GO Term size`), where

  + `GO_Term_rate` = n~unique_GO_Terms~ from `list of interest` / n~total~ `GO`s

  + `Gene_freq` = sum( P~r~ (each gene from `list of interest` arising in *any* `GO Term`) )

    - i.e. **how often does gene X get annotated to a different GO Term?**

  + `GO Term size` = n~total_genes~ in a `GO Term`

```{r Generate Overlap Probs, include = T, warning = F, message = F, echo = T}

  ## Initialize variables for generating the weighted overlap probabilities for each term / gene

  ## how frequently they occur across all GO terms
gene_freq = list()

  ## overlap variables

    ## (aka how many genes in the list of interest were detected in GO Term X? --> Overlap)
    ## for weighted:
      ## account for: 
        ## list size
        ## term size
        ## number of unique terms found from list of interest
        ## how frequently genes from the list are found in any GO terms
GO_INFO_by_TERM_df = GO_INFO_by_TERM_df %>%
  dplyr::mutate(Overlap_Prob = (Num_Genes_in_Term_from_List / max(Num_Genes_in_Term)
                                ),
                Weighted_Overlap_Prob = NA_real_)

  ## Populate
for (i in 1:length(list_of_interest)){
  gene_freq[[i]] = 
    (
      as.numeric(GENE_GO_INFO_df$Num_GO_Terms[which(GENE_GO_INFO_df$GeneID == list_of_interest[i])]) /
        42837
      )
}
names(gene_freq) = c(list_of_interest)

for (i in 1:nrow(GO_INFO_by_TERM_df) ) {
  for (j in 1:length(GO_INFO_by_TERM_df$Num_Genes_in_Term[i]) ) {
    
     GO_INFO_by_TERM_df$Weighted_Overlap_Prob[i] = 
       
       ## overlap rate relative to list size
       (GO_INFO_by_TERM_df$Num_Genes_in_Term_from_List[i]/as.numeric(length(list_of_interest))*
          ## gene annotation frequency
          as.numeric(unlist(gene_freq))[j]*
          ## GO Term frequency
          (as.numeric(length(Unique_GOs))/42837))/
       ## list size relative to the genes in the GO Term
       (as.numeric(length(list_of_interest))/GO_INFO_by_TERM_df$Num_Genes_in_Term[i])
  }
}
```

## Plot Probability Distributions

Plot the `Weighted Overlap Prob` distribution and (unweighted) `Overlap Prob` distribution

+ distributions of probabilities that given genes/proteins from a provided list are all identified within a given `GO Term`
+ `weighted` accounts for list size, term size, frequency of a gene occurring in any term

```{r Plot Overlap Probability Distributions, include = T, message = F, warning = F, echo = T, render = T}
  ## Inspect the probability distributions
GO_INFO_by_TERM_df %>%
  dplyr::select(GO_Term, Overlap_Prob, Weighted_Overlap_Prob) %>%
  pivot_longer(cols = c(contains("Overlap")),
               names_to = "Overlap Prob Type",
               values_to = "Prob") %>%
  ggplot(aes(x = Prob, fill = `Overlap Prob Type`)) +
  geom_histogram(aes(x = Prob))
```

Weighting broadens the binomial distribution of probabilities away from zero

# Hypothesis Test {.tabset .tabset-pills .tabset-fade}

Are the `Overlap Probabilities` significantly different than we would expect by
chance?

+ Perform a statistical test on each term's `Overlap Probability`:

  + binomial exact test compares:
    
    1. the probability of the number of genes identified from the provided list involved in a particular `GO Term` (`Overlap Prob`)
    2. the theoretical (`Overlap Prob`) probability of the number of genes involved in a particular `GO Term`
        
        - theoretical probability is determined by:
            - simulating 1000 draws from a binomial distribution for a `GO Term` of the same size with the probability determined from the provided list (above)
          - averaging those 1000 numbers of simulated "successes"
          - dividing by the maximum number of genes in all `GO Term`s
  
  + test returns a p-value of a two-sided binomial exact test

+ It is [known](https://geneontology.org/docs/faq/) that statistical results from this analysis may not be meaningful to the user because this approach will return large, vague `GO Terms` OR incredibly noisy and specific `GO Terms`, with a heavy dependence on the list size

## Binomial Exact Test 1

Perform an exact test across the derived probabilities for each `GO Term` to determine
if that term is over- or under- represented within the provided list:

-   `Weighted Overlap Probs` vs `Expected`

-   `Unweighted Overlap Probs` vs `Expected`

```{r Determine significance, include = T, message = F, warning = F, echo = T}
  ## Create variables for hypothetical probabilities, drawing from binomial distribution and performing an exact test
GO_INFO_by_TERM_df = GO_INFO_by_TERM_df %>%
  dplyr::mutate(E_Weighted_Overlap = NA_real_,
                E_Weighted_Overlap_Prob = NA_real_,
                Weighted_Overlap_Prob_p = NA_real_,
                E_Overlap = NA_real_,
                E_Overlap_Prob = NA_real_,
                Overlap_Prob_p = NA_real_)

  ## Fill variables
for( i in 1:nrow(GO_INFO_by_TERM_df) ) {
  
  ## what's the expected weighted overlap? probability? -> 1000 draws
    ## take the average of the 1000 draws for the overlap
      ## re-scale by factor of 10
  GO_INFO_by_TERM_df$E_Weighted_Overlap[i] = round(mean(
    rbinom(1000,
           size = GO_INFO_by_TERM_df$Num_Genes_in_Term[i],
           prob = GO_INFO_by_TERM_df$Weighted_Overlap_Prob[i]
           )
    ))
  
  ## same but un-weighted
  GO_INFO_by_TERM_df$E_Overlap[i] = round(mean(
    rbinom(n = 1000,
           size = GO_INFO_by_TERM_df$Num_Genes_in_Term[i],
           prob = GO_INFO_by_TERM_df$Overlap_Prob[i]
           )
    ))
}
  
    ## divide it by the total number of genes
GO_INFO_by_TERM_df$E_Weighted_Overlap_Prob = 
  
  GO_INFO_by_TERM_df$E_Weighted_Overlap / max(GO_INFO_by_TERM_df$Num_Genes_in_Term)


GO_INFO_by_TERM_df$E_Overlap_Prob = 
  
  GO_INFO_by_TERM_df$E_Overlap / max(GO_INFO_by_TERM_df$Num_Genes_in_Term)

for( i in 1:nrow(GO_INFO_by_TERM_df)) {
    ## perform an exact binomial test and extract p-values
  GO_INFO_by_TERM_df$Weighted_Overlap_Prob_p[i] =
    
    binom.test(GO_INFO_by_TERM_df$Num_Genes_in_Term_from_List[i],
               GO_INFO_by_TERM_df$Num_Genes_in_Term[i],
               (GO_INFO_by_TERM_df$E_Weighted_Overlap_Prob[i]),
               alternative = "two.sided")$p.value
  
  
  GO_INFO_by_TERM_df$Overlap_Prob_p[i] = 
    
    binom.test(GO_INFO_by_TERM_df$Num_Genes_in_Term_from_List[i],
                GO_INFO_by_TERM_df$Num_Genes_in_Term[i],
                (GO_INFO_by_TERM_df$E_Overlap_Prob[i]),
                alternative = "two.sided")$p.value
}
 
  ## any significant? remove low count (1 hit) data and add adjusted p-values
GO_INFO_by_TERM_df %>%
  dplyr::arrange(Weighted_Overlap_Prob) %>%
  dplyr::mutate(Weighted_Overlap_Prob_ADJP = (Weighted_Overlap_Prob_p / nrow(GO_INFO_by_TERM_df)) * 0.05,
                 Overlap_Prob_ADJP = (Overlap_Prob_p / nrow(GO_INFO_by_TERM_df)) * 0.05) %>%
  dplyr::filter(Num_Genes_in_Term_from_List > 1) %>%
  dplyr::slice(c(1:15)) %>%
  dplyr::select(c(GO_Term, Num_Genes_in_Term_from_List, Num_Genes_in_Term, contains("Overlap")))
```

Results are plausible, given what we suspect about this dataset

Sorting by p-value returns a flood of 0 p-values (unhelpful)

## Plot Expected Probs vs Actual Probs

Plot the `Expected Overlap Probabilities` (x axis) and the `Overlap Probabilities` (y axis)

  + `weighted overlap`

  + `unweighted overlap`

  + include a generalized additive model fit for each

```{r Prob plots 2, include = T, warning = F, message = F, echo = F, render = T, fig.width= 10}
  # plot probability distributions
GO_INFO_by_TERM_df = GO_INFO_by_TERM_df %>%
  dplyr::mutate(weighted_gam_preds = as.numeric(
                  predict.gam(
                    gam(data = GO_INFO_by_TERM_df,
                        formula = Weighted_Overlap_Prob ~ s(E_Weighted_Overlap_Prob,
                                                            k = 3,
                                                            bs = "cr"), weights = E_Weighted_Overlap_Prob
                        )
                    )
                  ),
                gam_preds = as.numeric(
                  predict.gam(
                    gam(data = GO_INFO_by_TERM_df,
                        formula = Overlap_Prob ~ s(E_Overlap_Prob,
                                                   k = 3,
                                                   bs = "cr"), weights = E_Overlap_Prob
                        )
                  )
                ),
                gam_fit_on_weighted_gam_preds = as.numeric(
                  predict.gam(
                    gam(data = GO_INFO_by_TERM_df,
                        formula = Weighted_Overlap_Prob ~ s(weighted_gam_preds,
                                                            k = 3,
                                                            bs = "cr"), weights = weighted_gam_preds
                    )
                  )
                ),
                gam_fit_on_unweighted_gam_preds = as.numeric(
                  predict.gam(
                    gam(data = GO_INFO_by_TERM_df,
                        formula = Overlap_Prob ~ s(gam_preds,
                                                   k = 3,
                                                   bs = "cr"), weights = gam_preds
                    )
                  )
                )
                )

weighted_fit_plot = GO_INFO_by_TERM_df %>%
  dplyr::select(c(contains("Weighted"), weighted_gam_preds, Num_Genes_in_Term_from_List)) %>%
  ggplot()+
  geom_point(aes(x = E_Weighted_Overlap_Prob,
                 y = Weighted_Overlap_Prob),
             color = "navy",
             alpha = 0.3) +
  geom_smooth(method = "gam",
              aes(x = `E_Weighted_Overlap_Prob`,
                  y = `weighted_gam_preds`),
              color = "black",
              linetype = "dashed") +
  labs(title = "Weighted Overlap Probabilities\nfor each unique GO Term from List")



unweighted_fit_plot = GO_INFO_by_TERM_df %>%
  dplyr::select(c(contains("Overlap"), gam_preds, Num_Genes_in_Term_from_List)) %>%
  ggplot()+
  geom_point(aes(x = E_Overlap_Prob,
                 y = Overlap_Prob),
             color = "navy",
             alpha = 0.3) +
  geom_smooth(method = "gam",
              aes(x = `E_Overlap_Prob`,
                  y = `gam_preds`),
              color = "black",
              linetype = "dashed") +
  labs(title = "Unweighted Overlap Probabilities\nfor each unique GO Term from List")

weighted_regressed_plot = GO_INFO_by_TERM_df %>%
  dplyr::select(c(contains("Weighted"), gam_preds, weighted_gam_preds, contains("fit_on"))) %>%
  ggplot() +
  geom_point(aes(x = E_Weighted_Overlap_Prob,
                 y = gam_fit_on_weighted_gam_preds),
             color = "darkgoldenrod4",
             alpha = 0.3) +
  geom_smooth(method = "gam",
              aes(x = `E_Weighted_Overlap_Prob`,
                  y = `gam_fit_on_weighted_gam_preds`),
              color = "black",
              linetype = "dashed") +
  labs(title = "Regressed Weighted Overlap Probabilities\nfor each unique GO Term from List",
       y = "Regressed_Weighted_Overlap_Prob")

unweighted_regressed_plot = GO_INFO_by_TERM_df %>%
  dplyr::select(c(contains("Overlap"), gam_preds, weighted_gam_preds, contains("fit_on"))) %>%
  ggplot() +
  geom_point(aes(x = E_Overlap_Prob,
                 y = gam_fit_on_unweighted_gam_preds),
             color = "darkgoldenrod4",
             alpha = 0.3) +
  geom_smooth(method = "gam",
              aes(x = `E_Overlap_Prob`,
                  y = `gam_fit_on_unweighted_gam_preds`),
              color = "black",
              linetype = "dashed") +
  labs(title = "Regressed Unweighted Overlap Probabilities\nfor each unique GO Term from List",
       y = "Regressed_Overlap_Prob")


ggarrange(weighted_fit_plot, unweighted_fit_plot, weighted_regressed_plot, unweighted_regressed_plot, ncol = 2, nrow = 2, common.legend = T)

```


## Assess GAM Model fits

How does the weighted overlap compare?

```{r Model fits, include = T, message = F, warning = F, echo = T}

GO_INFO_by_TERM_df %>%
  dplyr::mutate(
                Weighted_Overlap_RMSE = sqrt(mean(weighted_gam_preds - Weighted_Overlap_Prob)^2),
                
                Overlap_RMSE = sqrt(mean(gam_preds - Overlap_Prob)^2),
                
                Weighted_gam_Preds_RMSE = sqrt(mean(gam_fit_on_weighted_gam_preds - Weighted_Overlap_Prob)^2),

                Unweighted_gam_Preds_RMSE = sqrt(mean(gam_fit_on_unweighted_gam_preds - Overlap_Prob)^2),
                
                ) %>%
  dplyr::select(contains("RMSE")) %>%
  dplyr::distinct() %>%
  pivot_longer(cols = c(colnames(.)))
```

All fits are comparable

*We'll use the* `regressed weighted GAM probabilities` in the `binomial exact test`

## Binomial Exact Test 2

Re-compute the binomial exact test using the `regressed weighted GAM fit values` as the the `Weighted Overlap Probabilities`

i.e. Regress the `Weighted Overlap Prob`s and use those regressed values as the theoretical probs in a `binomial exact test`


```{r Re-compute binomial exact test probabilities, include = T, message = F, warning = F, echo = T}

GO_INFO_by_TERM_df = GO_INFO_by_TERM_df %>%
  dplyr::mutate(
    E_Weighted_Overlap_2 = NA_real_,
    E_Weighted_Overlap_Prob_2 = NA_real_,
    Weighted_Overlap_Prob_p_2 = NA_real_,
    Weighted_Overlap_Prob_ADJP_2 = NA_real_,
    E_Overlap_2 = NA_real_,
    E_Overlap_Prob_2 = NA_real_,
    Overlap_Prob_p_2 = NA_real_,
    Overlap_Prob_ADJP_2 = NA_real_
  )

for( i in 1:nrow(GO_INFO_by_TERM_df) ) {
  
  ## what's the expected weighted overlap? probability? -> 1000 draws; USE THE GAM PROB PREDICTS
    ## take the average of the 1000 draws for the overlap
  GO_INFO_by_TERM_df$E_Weighted_Overlap_2[i] = round(mean(
    rbinom(1000,
           size = GO_INFO_by_TERM_df$Num_Genes_in_Term[i],
           prob = GO_INFO_by_TERM_df$gam_fit_on_weighted_gam_preds[i]
           )
    )
  )
  
  GO_INFO_by_TERM_df$E_Overlap_2[i] = round(mean(
    rbinom(1000,
           size = GO_INFO_by_TERM_df$Num_Genes_in_Term[i],
           prob = GO_INFO_by_TERM_df$gam_fit_on_unweighted_gam_preds[i]
           )
  )
  )
}
  
    ## divide predicted overlap counts by total number of genes
GO_INFO_by_TERM_df$E_Weighted_Overlap_Prob_2 = 
  
  GO_INFO_by_TERM_df$E_Weighted_Overlap_2 / max(GO_INFO_by_TERM_df$Num_Genes_in_Term)


GO_INFO_by_TERM_df$E_Overlap_Prob_2 = 
  
  GO_INFO_by_TERM_df$E_Overlap_2 / max(GO_INFO_by_TERM_df$Num_Genes_in_Term)

    ## perform an exact binomial test and extract p-values
for( i in 1:nrow(GO_INFO_by_TERM_df) ) {
  
  GO_INFO_by_TERM_df$Weighted_Overlap_Prob_p_2[i] = 
    
    binom.test(GO_INFO_by_TERM_df$Num_Genes_in_Term_from_List[i],
               GO_INFO_by_TERM_df$Num_Genes_in_Term[i],
               (GO_INFO_by_TERM_df$E_Weighted_Overlap_Prob_2[i]),
               alternative = "two.sided")$p.value
  
  GO_INFO_by_TERM_df$Overlap_Prob_p_2[i] = 
    
    binom.test(GO_INFO_by_TERM_df$Num_Genes_in_Term_from_List[i],
               GO_INFO_by_TERM_df$Num_Genes_in_Term[i],
               (GO_INFO_by_TERM_df$E_Overlap_Prob_2[i]),
               alternative = "two.sided")$p.value
  
}


GO_INFO_by_TERM_df$Weighted_Overlap_Prob_ADJP_2 = 
  
  (GO_INFO_by_TERM_df$Weighted_Overlap_Prob_p_2 / nrow(GO_INFO_by_TERM_df) ) * 0.05


GO_INFO_by_TERM_df$Overlap_Prob_ADJP_2 = 
  
  (GO_INFO_by_TERM_df$Overlap_Prob_p_2 / nrow(GO_INFO_by_TERM_df) ) * 0.05
  
```

## Do Exact Test Results Confirm?

```{r GO Bar Plot, include = T, message = F, warning = F, echo = F, render = T}
GO_Bar = GO_INFO_by_TERM_df %>%
  dplyr::arrange(Weighted_Overlap_Prob) %>%
  dplyr::filter(between(Num_Genes_in_Term_from_List, 2, round(as.numeric(length(list_of_interest))/2))) %>%
  dplyr::slice(c(1:15)) %>%
  dplyr::select(c(GO_Term,
                  
                  Num_Genes_in_Term_from_List,
                  
                  E_Weighted_Overlap,
                  
                  E_Weighted_Overlap_2,
                  
                  Num_Genes_in_Term,
                  
                  Weighted_Overlap_Prob,
                  
                  E_Weighted_Overlap_Prob_2,
                  
                  Weighted_Overlap_Prob_p,
                  
                  Weighted_Overlap_Prob_p_2,
                  
                  Weighted_Overlap_Prob_ADJP_2,
                  
                  gam_fit_on_weighted_gam_preds))

GO_Bar = GO_Bar %>% 
  dplyr::mutate(GO_Term = case_when(grepl(GO_Term, pattern = "V0 domain") ~
                                        "vacuolar protein-transporting\nV-type ATPase, V0 domain",
                                      grepl(GO_Term, pattern = "retrograde transport") ~
                                        "retrograde transport,\nvesicle recycling within Golgi",
                                      grepl(GO_Term, pattern = "ubiquitin") ~
                                        "ubiquitin-independent protein\ncatabolic process via the multivesicular\nbody sorting pathway",
                                      grepl(GO_Term, pattern = "negative regulation") ~
                                        "negative regulation of\nmammary gland epithelial cell proliferation",
                                      grepl(GO_Term, pattern = "late endosome") ~
                                        "late endosome to lysosome transport\nvia multivesicular body sorting pathway",
                                      TRUE ~ GO_Term))
GO_Bar %>%
  ggplot(aes(x = reorder(GO_Bar$GO_Term, GO_Bar$Weighted_Overlap_Prob),
             y = GO_Bar$Weighted_Overlap_Prob)) +
  geom_col(stat = `identity`, aes(fill = `Weighted_Overlap_Prob`))+
  scale_fill_continuous(name = "Weighted\nOverlap\nProbability")+
  labs(x = "GO Term",
       y = "Weighted Overlap Prob",
       title = "Estimated Probabilities of\nGenes from Given List Occurring in Given GO Term")+
  theme_light() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5),
        axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0)),
        plot.title = element_text(hjust = 1),
        panel.grid.major.y = element_blank()) +
  coord_flip()
```

```{r Results Sorted by Weighted Overlap Probability, include = T, message = F, warning = F, echo = F}
GO_INFO_by_TERM_df %>%
  dplyr::arrange(Weighted_Overlap_Prob) %>%
  dplyr::filter(between(Num_Genes_in_Term_from_List, 2, round(as.numeric(length(list_of_interest))/2))) %>%
  dplyr::slice(c(1:15)) %>%
  dplyr::select(c(GO_Term,
                  
                  Num_Genes_in_Term_from_List,
                  
                  Num_Genes_in_Term,
                  
                  Weighted_Overlap_Prob,
                  
                  weighted_gam_preds,
                  
                  gam_fit_on_weighted_gam_preds,
                  
                  E_Weighted_Overlap_Prob_2,
                  
                  Weighted_Overlap_Prob_ADJP_2))
```

# Conclusions

**No, test results don't confirm because values are too small to determine if probabilities or counts are significantly different than by chance.**

But when removing highly specific `GO Terms` (with only 1 gene) and arranging by the least probable weighted overlap, the results are incredibly plausible:

`GO Term`s we suspect to be intimately and robustly releated to the target protein are *incredibly improbable* to be associated with the target protein by chance

```{r Export Results, include = F, warning = F, message = F, echo = F}

# write.csv("C:/Users/Erik/Desktop/BoxCopy/Lab/Omics/MS/Bioinformatics Analysis Output/6_15_2021/IntersectList.csv")
# save.image("C:/Users/Erik/Desktop/BoxCopy/Lab/Omics/MS/Bioinformatics Analysis Output/6_15_2021/proteomics_GO_INFO.RData")
```
